---
title: "p8105_hw5_cw3747"
author: "chuhan wang_cw3747"
date: "2025-11-04"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(patchwork)
library(scales)
library(dplyr)


knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Birthday Function
```{r}
library(ggplot2)
library(dplyr)

set.seed(123)


birthday_dup <- function(n, days = 365) {
  bdays <- sample.int(days, n, replace = TRUE) 
  any(duplicated(bdays))                       
}
```


### iteration
```{r}
simulate_prob <- function(n, B = 10000, days = 365) {
  mean(replicate(B, birthday_dup(n, days)))
}

ns <- 2:50
B  <- 10000

probs <- numeric(length(ns))
for (i in seq_along(ns)) {
  probs[i] <- simulate_prob(ns[i], B = B)
}

df_loop <- data.frame(n = ns, p = probs)

n50_loop <- min(df_loop$n[df_loop$p >= 0.5])
```

### ggplot
```{r}
ggplot(df_loop, aes(n, p)) +
  geom_line(linewidth = 1) +
  geom_point(size = 1.5) +
  geom_hline(yintercept = 0.5, linetype = 2) +
  geom_vline(xintercept = n50_loop, linetype = 2) +
  annotate("text", x = n50_loop + 1, y = 0.52,
           label = paste0("~50% at n = ", n50_loop), hjust = 0) +
  labs(title = "P(at least one shared birthday) vs. Group size",
       x = "Group size (n)", y = "Probability") +
  theme_minimal(base_size = 12)
```

The plot shows that the probability of at least one shared birthday increases rapidly with group size. Around n = 23, the probability reaches about 0.5, meaning that in a random group of 23 people there is 50% chance that some pair shares a birthday (this does not mean that “half of the people” have a match). As n approaches 50, the probability nears 1, implying that a shared-birthday pair is almost guaranteed, though not that everyone finds a match. Real-world seasonality and clustering can shift the curve slightly, but the qualitative “birthday paradox” pattern remains.


