---
title: "p8105_hw5_cw3747"
author: "chuhan wang_cw3747"
date: "2025-11-04"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(patchwork)
library(scales)
library(dplyr)


knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem1

### Birthday Function
```{r}
library(ggplot2)
library(dplyr)

set.seed(123)


birthday_dup <- function(n, days = 365) {
  bdays <- sample.int(days, n, replace = TRUE) 
  any(duplicated(bdays))                       
}
```


### iteration
```{r}
simulate_prob <- function(n, B = 10000, days = 365) {
  mean(replicate(B, birthday_dup(n, days)))
}

ns <- 2:50
B  <- 10000

probs <- numeric(length(ns))
for (i in seq_along(ns)) {
  probs[i] <- simulate_prob(ns[i], B = B)
}

df_loop <- data.frame(n = ns, p = probs)

n50_loop <- min(df_loop$n[df_loop$p >= 0.5])
```

### ggplot
```{r}
ggplot(df_loop, aes(n, p)) +
  geom_line(linewidth = 1) +
  geom_point(size = 1.5) +
  geom_hline(yintercept = 0.5, linetype = 2) +
  geom_vline(xintercept = n50_loop, linetype = 2) +
  annotate("text", x = n50_loop + 1, y = 0.52,
           label = paste0("~50% at n = ", n50_loop), hjust = 0) +
  labs(title = "P(at least one shared birthday) vs. Group size",
       x = "Group size (n)", y = "Probability") +
  theme_minimal(base_size = 12)
```

The plot shows that the probability of at least one shared birthday increases rapidly with group size. Around n = 23, the probability reaches about 0.5, meaning that in a random group of 23 people there is 50% chance that some pair shares a birthday (this does not mean that “half of the people” have a match). As n approaches 50, the probability nears 1, implying that a shared-birthday pair is almost guaranteed, though not that everyone finds a match. Real-world seasonality and clustering can shift the curve slightly, but the qualitative “birthday paradox” pattern remains.


## Problem2

### set function
```{r}
library(broom) 

set.seed(123)

sim_once <- function(mu, n = 30, sigma = 5) {
  x  <- rnorm(n, mean = mu, sd = sigma)
  tt <- t.test(x, mu = 0)
  tibble(
    mu = mu,
    mu_hat = mean(x),
    p = tidy(tt)$p.value,
    reject = p < 0.05
  )
}
```

### iteration
```{r}
simulate_mu <- function(mu, B = 5000, n = 30, sigma = 5) {
  bind_rows(replicate(B, sim_once(mu, n, sigma), simplify = FALSE))
}
```

### for different $\mu$
```{r}
mus <- 0:6
res_list <- vector("list", length(mus))
for (i in seq_along(mus)) {
  res_list[[i]] <- simulate_mu(mus[i], B = 5000)
}
res <- bind_rows(res_list)
```

### power of test
```{r}
summ <- res |>
  group_by(mu) |>
  summarise(
    power = mean(reject),
    avg_mu_hat_all = mean(mu_hat),
    avg_mu_hat_rej = mean(mu_hat[reject]),
    .groups = "drop"
  )

ggplot(summ, aes(mu, power)) +
  geom_line(linewidth = 1) +
  geom_point(size = 1.6) +
  geom_hline(yintercept = 0.05, linetype = 2) +
  labs(title = "Power of t-test",
       x = "True mean (μ)", y = "Power") +
  theme_minimal(base_size = 12)
```

We fix $\mu_0 = 0$ and $\sigma = 5$. The effect size is $d = (\mu - \mu_0) / \sigma$. Hence, larger $\mu$ implies a larger effect size. The plot shows that power increases monotonically with $\mu$: at $\mu$ = 0, the the rejection rate is about the rejection rate is about $\alpha$ = 0.05; as $\mu$ grows, power rises quickly. Larger effect size makes the sample mean more distinct from 0, yielding a larger absolute t-statistic and a higher chance of rejecting the null.

